---
title: "R Notebook"
output: html_notebook
---

Thank you for contributing to covidregionaldata! This tutorial will take you through adding a new subnational dataset.

How to use this tutorial:
- Our datasets are implemented using R6 methods. If you are already very familiar with these, this guide may be too basic for you.
- First, clone the covidregionaldata repository and create a new branch. 
- Copy the [country template R script](./R/country-template.R), rename it to the name of the country you would like to add, and save into the R/ folder.
- This guide will take you through how to add and clean a dataset using this script. Because we use R6 methods, this can be difficult to see in the script itself. Please use this notebook first and copy across to the template where indicated.
- You will also need to add tests.
- When finished, please open a pull request. Our team will review the code and work with you to make any changes needed before we can merge the dataset into the package.
- Where possible, we use the tidyverse suite of packages for manipulating data.

```{r}
library(dplyr)
```

### Linking a dataset

Add the url for the new dataset.

```{r}
data_url <- "https://github.com/wcota/covid19br/raw/master/cases-brazil-cities-time.csv.gz"
```

