---
title: "How to add a data set"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to add a data set}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(covidregionaldata)
```

## Recipe for adding a new data set for `covidregionaldata`

### Ingredients

You need an open and accessible data source, preferably in the form of 
a CSV file updated on a regular basis and accessible for download with a
fixed (or programmably-predictable) URL.

`covidregionaldata` provides an internal function, `csv_reader` which 
does nearly all the hard work of downloading, caching and parsing such a
CSV. 

If you can, you may wish to download a local copy of your data source's
CSV to work with. `covidregionaldata` internally uses [`vroom`](https://vroom.r-lib.org/index.html) to parse most CSV files.

Run `vroom` on your local file, see what you get.

## Create a new file with the name of your country

This is the main working portion of your addition. It should provide at
least one, maybe two, functions which download data from your data source,
process it minimally and rename some columns where necessary so that other
parts of the package can 

### Add in `get_country_regional_cases_only_level_1` functions

### Add in `get_country_regional_cases_with_level_2` functions

## Add your new `get_` functions to get_regional_data

## Add to get_info_covidregionaldata

## Add to get_region_codes.R

* Add in the local name of your level 1 administrative units
* Add in the local name of your level 2 administrative units (if applicable)
* Provide a function which provides a look-up table for level 1 administrative units
  * tribble is useful for this, as is datapasta
* Provide a function which provides a look-up table for level 2 administrative units (if applicable)

## Add a line to README.md

## Edit DESCRIPTION

## Edit WORDLIST

