---
title: "Package overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Installing and loading the package

The package can either be installed from CRAN, from our `drat` repository, or from GitHub. See the README for details. Once installed load the package using the following,

```{r}
library(covidregionaldata)
```

## Worldwide data

### Accessing national data

Both the World Health Organisation (WHO) and European Centre for Disease Control (ECDC) provide worldwide national data. Access national level data for any country using:

```{r}
get_national_data()
```

This returns daily new and cumulative (total) cases, and where available, deaths, hospitalisations, and tests. For a complete list of variables returned, see section 5, "Data glossary" below. See the documentation (`?get_national_data`) for details of optional arguments.

Data is returned with no gaps in the structure of the data by country over time, and NAs fill in where data are not available.

## Sub-national time-series data

### Accessing sub-national data

Access sub-national level data for a specific country over time using `get_regional_data()`. Use `get_available_datasets()` to explore the currently supported sub-national datasets and select the data set of interest using the `country` (selects the country of interest), and `level` (selects the spatial scale of the data) arguments of `get_regional_data`.

This function returns daily new and cumulative (total) cases, and where available, deaths, hospitalisations, and tests. For a complete list of variables returned, see section 5, "Data glossary" below. See the documentation (`?get_regional_data`) for details of optional arguments.

As for national level data any gaps in reported data are filled with NAs.

For example, data for Belgium Level 1 regions over time can be accessed using:

```{r}
get_regional_data(countries = "Belgium")
```
This returns a dataset in this format:

**date**|**region**|**iso\_code**|**cases\_new**|**cases\_total**|**deaths\_new**|**deaths\_total**|**recovered\_new**|**recovered\_total**|**hosp\_new**|**hosp\_total**|**tested\_new**|**tested\_total**
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
2020-05-24|Wallonia|BE-WAL|24|18196|16|3251|NA|NA|8|5126|NA|NA
2020-05-25|Brussels|BE-BRU|26|5838|2|1421|NA|NA|6|2533|NA|NA
2020-05-25|Flanders|BE-VLG|183|32381|14|4681|NA|NA|29|9334|NA|NA

Alternatively, the same data can be accessed using the underlying class as follows (the belgium object now contains data at each processing step and the methods used at each step), 

```{r}
belgium <- Belgium$new(get = TRUE)
belgium$return()
```
### Level 1 and Level 2 regions

All countries included in the package (see below,"Coverage") have data for regions at the admin-1 level, the largest administrative unit of the country (e.g. state in the USA). Some countries also have data for smaller areas at the admin-2 level (e.g. county in the USA).

Data for Level 2 units can be returned by using the `level = "2"` argument. The dataset will still show the corresponding Level 1 region.

An example of a country with Level 2 units is Belgium, where Level 2 units are Belgian provinces:

```{r}
get_regional_data("Belgium", level = "2")
```
This returns a dataset with the format:

**date**|**province**|**level\_2\_region\_code**|**region**|**iso\_code**|**cases\_new**|**cases\_total**|**deaths\_new**|**deaths\_total**|**recovered\_new**|**recovered\_total**|**hosp\_new**|**hosp\_total**|**tested\_new**|**tested\_total**
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
2020-05-24|Brussels|BE-BRU|Brussels|BE-BRU|7|5812|NA|NA|NA|NA|4|2527|NA|NA
2020-05-24|Antwerpen|BE-VAN|Flanders|BE-VLG|16|7905|NA|NA|NA|NA|5|2510|NA|NA
2020-05-24|Limburg|BE-VLI|Flanders|BE-VLG|14|6126|NA|NA|NA|NA|2|1848|NA|NA

### Totals

For totalled data up to the most recent date available, use the `totals` argument.

```{r}
get_regional_data("Belgium", totals = TRUE)
```
This returns a dataset with one row for each region, in the same format:

**region**|**iso\_code**|**cases\_total**|**deaths\_total**|**recovered\_total**|**hosp\_total**|**tested\_total**
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
Flanders|BE-VLG|34195|4878|0|9694|0
Wallonia|BE-WAL|19093|3362|0|5321|0
Brussels|BE-BRU|6229|1482|0|2657|0

## Data glossary

#### Subnational data

The data columns that will be returned by `get_regional_data()` are listed below.

To standardise across countries and regions, the columns returned for each country will _always_ be the same. If the corresponding data was missing from the original source then that data field is filled with NA values (or 0 if accessing totals data).

Note that Date is not included if the `totals` argument is set to TRUE. Level 2 region/level 2 region code are not included if the `level = "1"`.

* `date`: the date that the counts were reported (YYYY-MM-DD).

* `level_1_region`: the level 1 region name. This column will be named differently for different countries (e.g. state, province).

* `level_1_region_code`: a standard code for the level 1 region. The column name reflects the specific administrative code used. Typically data returns the iso_3166_2 standard, although where not available the column will be named differently to reflect its source.

* `level_2_region`: the level 2 region name. This column will be named differently for different countries (e.g. city, county).

* `level_2_region_code`: a standard code for the level 2 region. The column will be named differently for different countries (e.g. `fips` in the USA).

* `cases_new`: new reported cases for that day.

* `cases_total`: total reported cases up to and including that day.

* `deaths_new`: new reported deaths for that day.

* `deaths_total`: total reported deaths up to and including that day.

* `recovered_new`: new reported recoveries for that day.

* `recovered_total`: total reported recoveries up to and including that day.

* `hosp_new`: new reported hospitalisations for that day.

* `hosp_total`: total reported hospitalisations up to and including that day (note this is cumulative total of new reported, _not_ total currently in hospital).

* `tested_new`: tests for that day.

* `tested_total`: total tests completed up to and including that day.

#### National data

In addition to the above, the following columns are included when using `get_national_data()`.

* `un_region`: country geographical region defined by the United Nations.

* `who_region`: only included when `source = "WHO"`. Country geographical region defined by WHO.

* `population_2019`: only included when `source = "ECDC"`. Total country population estimate in 2019.
